---
title: "Regression models"
author: "Davyd"
date: "May, 2018"
output: pdf_document
---

## Executive Summary

You work for Motor Trend, a magazine about the automobile industry. Looking at a data set of a collection of cars, they are interested in exploring the relationship between a set of variables and miles per gallon (MPG) (outcome). They are particularly interested in the following two questions:

"Is an automatic or manual transmission better for MPG"
"Quantify the MPG difference between automatic and manual transmissions"

## Load Data

Load required packages, dataset, and convert categorical variables to factors.

```{r results='hide', message=FALSE}
library(ggplot2)
data(mtcars)
head(mtcars, n=3)
dim(mtcars)
mtcars$cyl <- as.factor(mtcars$cyl)
mtcars$vs <- as.factor(mtcars$vs)
mtcars$am <- factor(mtcars$am)
mtcars$gear <- factor(mtcars$gear)
mtcars$carb <- factor(mtcars$carb)
attach(mtcars)
```

## Exploratory Data Analysis

**See Appendix I** Box plot comparing MPG between Automatic and Manual 
transmission. The results show a significant  increase in MPG for manual transmission 
when compared to automatic transmission.

### Statistical Inference

T-Test between transmission type and MPG

```{r}
testResults <- t.test(mpg ~ am)
testResults$p.value
```

The T-Test rejects the null hypothesis that there is no difference in MPG for both transmission types.

```{r}
testResults$estimate
```

The estimated difference between the two transmission types is 7.24494 MPG in favour of manual transmission.

### Regression Analysis

Fitting the model for the data

```{r results='hide'}
fullModelFit <- lm(mpg ~ ., data = mtcars)
summary(fullModelFit)  # results hidden
summary(fullModelFit)$coeff  # results hidden
```

Since none of the p-values are below 0.05, we cannot conlude that there is any statistical significance.  

Selecting variables which are most statistically significant

```{r results='hide'}
stepFit <- step(fullModelFit)
summary(stepFit) # results hidden
summary(stepFit)$coeff # results hidden
```

The new model has 4 variables (cylinders, horsepower, weight, transmission). The
R-squared value of 0.8659 confirms that this model explains about 87% of the 
variance in MPG. The p-values also are statistically significantly because they
have a p-value less than 0.05. The coefficients conclude that increasing the 
number of cylinders from 4 to 6 with decrease the MPG by 3.03.  Further 
increasing the cylinders to 8 with decrease the MPG by 2.16.  Increasing the 
horsepower is decreases MPG 3.21 for every 100 horsepower.  Weight decreases the
MPG by 2.5 for each 1000 lbs increase. A Manual transmission improves the MPG by
1.81.

## Residuals & Diagnostics

Residual Plot

**See Appendix Figure II**

The plots show that:

1. The randomness of the Residuals vs. Fitted plot supports the assumption of
independence
2. The points of the Normal Q-Q plot following closely to the line conclude that
the distribution of residuals is normal
3. The Scale-Location plot random distribution confirms the constant variance 
assumption
4. Since all points are within the 0.05 lines, the Residuals vs. Leverage 
concludes that there are no outliers

```{r}
sum((abs(dfbetas(stepFit)))>1)
```

## Conclusion

There is a difference in MPG between transmission type. A manual transmission
will have a slight advantage in MPG. However, weight, horsepower, & 
number of cylinders are more statistically significant when determining MPG.

## Appendix Figures

## I

```{r echo=FALSE}
  boxplot(mpg ~ am, 
          xlab="Transmission Type (0 = Automatic, 1 = Manual)", 
          ylab="MPG",
          main="MPG by Transmission Type")
```

## II

```{r echo=FALSE}
par(mfrow = c(2, 2))
plot(stepFit)
```